---
title: "Lab1"
format: html
editor: visual
---

# Lab 1 - Deterministic Models and Probability I

```{r}
library(boot) # for inv.logit()
```

## II. Deterministic Models

### Plotting Deterministic Models

```{r}
par(mfrow = c(2,2))

# linear plot
linear = function(B0, B1, x){  #pay close attention to the syntax of the function command.  The stuff in parantheses are the inputs ot the function
  y = B0 + B1*x     #This is what the function does
  return(y)        #These are the values that the function returns
}
x = -100:100; B0 = 2; B1 = 0.5
y = linear(B0, B1, x)
plot(x,y,type = 'l', main = "linear")

# inv logit plot
inverse = function(B0, B1, x){
  y = inv.logit(B0 + B1*x)
  return(y)        #These are the values that the function returns
}
x = -100:100; B0 = 2; B1 = 0.5
y = inverse(B0, B1, x)
plot(x,y,type = 'l', main = "Inverse Logit")

# exponential plot
exponential = function(B0, B1, x){
  y = exp(B0 + B1*x)
  return(y)        #These are the values that the function returns
}
x = -100:200; B0 = 2; B1 = 0.5
y = exponential(B0, B1, x)
plot(x,y,type = 'l', main = "Exponential")

# aymptotic plot
asymptotic = function(B0, B1, x){
  y = (B0 * x)/(B0/2 + x)
  return(y)        #These are the values that the function returns
}
x = -100:100; B0 = 2; B1 = 0.5
y = asymptotic(B0, B1, x)
plot(x,y,type = 'l', main = "Asymptotic")
```

### Examples for Each Model

sociological or ecological

#### Linear

-   eco - tree growth and precipitation

-   socio - lifespan and wealth

#### Inverse Logit

-   eco - Abundance of a competitive invasive species after the introduction to a new area

-   eco - human growth rate, beacuse speeds way up at adolescence then slows down again

#### Exponential

-   eco - amount of carbon in the atmosphere and the greenhouse effect

-   socio - compounding interest

#### Asymptotic

-   a black hole and the light entering

-   eco - recovery of a formerly increasing species after a major stochastic event that nearly wiped them out

## III. The chain rule of probability

1\. Use the chain rule of probability to factor the joint distribution Pr(ğœƒ1,ğœƒ2,ğœƒ3,ğœƒ4)

into a joint conditional distribution.

Pr(ğœƒ1,ğœƒ2,ğœƒ3,ğœƒ4) = Pr(ğœƒ4\|ğœƒ3,ğœƒ2,ğœƒ1) \* Pr(ğœƒ3\|ğœƒ2,ğœƒ1) \* Pr(ğœƒ2\|ğœƒ1) \* Pr(ğœƒ1)

2\. Consider the following factored joint distribution:

Pr(ğœƒ,ğ‘¤,ğœ‚,ğ›¼)=Pr(ğ›¼âˆ£ğ‘¤,ğœ‚,ğœƒ)Pr(ğ‘¤âˆ£ğœ‚,ğœƒ)Pr(ğœ‚âˆ£ğœƒ)Pr(ğœƒ)

Simplify this equation using the knowledge that ğœ‚ and ğœƒ are independent and that ğ‘¤ and ğœƒ are independent.

Pr(ğœƒ,ğ‘¤,ğœ‚,ğ›¼) = Pr(ğ›¼âˆ£ğ‘¤,ğœ‚,ğœƒ) \* Pr(ğ‘¤âˆ£ğœ‚) \* Pr(ğœ‚) \* Pr(ğœƒ)
